Case study validations of automatic bullet matching

### 2018-10-25 index.Rmd
table expanded
phrasing

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope and visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling). In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.

## 3D Imaging Technology
3D characterization of bullet surfaces has been considered since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option. More recently, approaches using 3D measurement data were proposed starting in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}. These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques. In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements. This allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.

Commonly, approaches derived from 3D surface measurements use some features which are developed to mimic the perception of FTEs\citep{luAutomatedBulletIdentification2014}. Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion. Automated approaches to estimation of width of the land engraved area and twist angle were described in \citet{chuPilotStudyAutomated2010}. Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements; in some cases, these features mimic features used by FTEs during image analysis. One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}. Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae \citep{chuAutomaticIdentificationBullet2013}.

In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs). From each land engraved area a signature is derived by

>>>>

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope. They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling). In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.

## 3D Imaging Technology
More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option. \svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}. More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}. These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques. In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements. Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks. \svp{In addition, there is significantly more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}

Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those qualitatively analyzed by FTEs}\citep{luAutomatedBulletIdentification2014}. Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion. Automated approaches to estimation of width of the land engraved area and twist angle were described in \citet{chuPilotStudyAutomated2010}. Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements. One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}. Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae \citep{chuAutomaticIdentificationBullet2013}.

In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest\citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs). From each land engraved area a signature is derived by
